---
title: "Interactive"
author: "Valerie Chavez"
date: "2022-09-27"
output:
  pdf_document: default
  html_document: default
---
---
title: "Interactive"
output: html_notebook
---

I wish to explore secondary student responses to the questions: 
- Overall, how do you feel about your life? 
- How happy have you been feeling this week?

And compare difference across race/ethnicity and gender
Responses were on a 5 point likert scale, which consisted of 5 emojis ranging from very sad to very happy for the first question. And responses for the second question rangd from never happy (1) to happy all the time (5).

Think about your life lately. How happy have you been feeling this week?
m_em_feel_happy pos


**Data is provided by and property of Youth Truth Student Survey, a national nonprofit, and may only be shared in aggregate for the confidentiality of students and clients.**

Our sample consisted of 161,340 secondary students (Grades 6-12) in the 2021-22 school year across 19 states, and 442 schools.(Note: not all students reponded to every question, so the sample is closer to 130,000).

For the confidentiallity of clients and students, I can't share the original data. 

Schools that choose to work with Youth Truth, and opt in to the Emotional and Mental health additional topic administered the questions to students. 


Loading various libraries (as I was testing I lost track of which I used and chose not to use, so I kept them all in!)
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(plotly)
library(htmlwidgets)
library(stringr)

```

Loading and prepping data
```{r cars}
HS<-read.csv("/Users/valerier/Dropbox (CEP)/YouthTruth/Data and Research/EMH Back to School 2022/R Script and Results/HS/HS_dataclean_2022.csv")
MS<-read.csv("/Users/valerier/Dropbox (CEP)/YouthTruth/Data and Research/EMH Back to School 2022/R Script and Results/MS/MS_dataclean_2022.csv")

```


```{r}
HS_subset<- HS[ ,c("em_life","gender", "racen", "em_feel_happy")]

MS_subset<-MS[ ,c("m_em_life","m_gender", "m_racen", "m_em_feel_happy")]

colnames(MS_subset)<-c("em_life","gender", "racen", "em_feel_happy")

Secondary<-rbind(HS_subset,MS_subset)

Secondary<-na.omit(Secondary)

summary(Secondary)
table(Secondary$gender)
table(Secondary$race)

```

```{r}
#removing n/as, skips, prefer not to say. etc. 

Secondary<-Secondary %>% filter(gender != 77 & gender !=99)
Secondary<-Secondary %>% filter(racen != 77 & racen !=99 & racen !=9)
table(Secondary$gender)
table(Secondary$racen)

dim(Secondary)

```



```{r}
#replacing gender codes with names so it is easier to highlight

Secondary<-Secondary %>%
            mutate(gender=replace(gender,gender==1, "Boy/Man"))
Secondary<-Secondary %>%
            mutate(gender=replace(gender,gender==2, "Girl/Woman"))
Secondary<-Secondary %>%
            mutate(gender=replace(gender,gender==3, "Non-Binary"))




```

```{r}
#Manually add our YT colors for our pallette
YTPalette<-c("#245971", "#f99c25","#b0c5cc")

set.seed(42)
EMH <-ggplot(Secondary, aes(x=em_life, y=em_feel_happy, group= gender, color=factor(gender)))+
  geom_count(alpha=.7, 
             position=position_jitterdodge(jitter.width = 0 , 
                                           jitter.height = .2, 
                                           dodge.width = .2), 
             aes(size=after_stat(prop), group = factor(gender)))+
  scale_size(range=c(2,12))+
  xlab(str_wrap("Overall, how do you feel about your life?")) + 
  ylab(str_wrap("How happy have you been feeling this week?")) +
  labs(title="Secondary Student Mental Health", colour= "Gender", shape= "Proportion of students by gender", caption = "Source:YouthTruth:youthtruthsurvey.org")+
  coord_equal()+
  scale_fill_manual(values= YTPalette,labels = c("Boy/Man", "Girl/Woman","Non-Binary"),aesthetics = "colour")+
  guides(colour = guide_legend(override.aes = list(size=5)), size = guide_legend("Prop", override.aes=list(col="#0e3051")))+
  theme_bw()+
  theme(plot.title = element_text(family = "Helvetica", colour = "#0e3051", face = "bold", size = (15)),
        legend.title = element_text(colour = "#245971", face = "bold", family = "Helvetica", , size = (10)),
        legend.text = element_text(face = "italic", colour = "#0e3051", family = "Helvetica"),
        axis.title = element_text (family = "Helvetica", size = (12), colour = "#245971"),
        axis.text = element_text(family = "Courier", colour = "#245971", size = (12)))




EMH

```



```{r}

EMHplotly<-ggplotly(EMH, width = 512, height = 512, tooltip= c("group", "prop"))%>% 
  layout(showlegend = FALSE)%>%
  highlight(on = "plotly_hover")

  
 
 EMHplotly
```


```{r}
saveWidget(as_widget(EMHplotly), "EMHplotly.html")

save(EMHplotly,file="EMHplotly.rda")
```


